### 2017-06-18 sec-conclusion.tex
Incorporate new citation from the editor.
Also shuffle around some figures to make the page layout prettier with
the now longer related work and model "derivation" sections.

Currently, the model learns a transition matrix over the latent states that
is \emph{shared} across all students. It would be interesting to instead
learn a different latent state transition matrix for each individual
student, but keep the second-level Markov models shared. This would provide
the model with more flexibility which may be desirable itself, but would
also naturally result in a description of a student (via his or her HMM
transitions) that could be incorporated into existing supervised learning
techniques that try to predict student outcomes for understanding
which of the latent behavior patterns discovered by 2L-HMM are most predictive
of the performance of student learning.

>>>>

Currently, the model learns a transition matrix over the latent states that
is \emph{shared} across all students. It would be interesting to instead
learn a different latent state transition matrix for each individual
student, but keep the second-level Markov models shared. This would provide
the model with more flexibility which may be desirable itself, but would
also naturally result in a description of a student (via his or her HMM
transitions) that could be incorporated into existing supervised learning
techniques that try to predict student outcomes for understanding
which of the latent behavior patterns discovered by 2L-HMM are most predictive
of the performance of student learning. One could also relax this somewhat
and posit that \emph{groups} of students transition between the lower layer
patterns our method identifies in distinct ways and instead do a soft
clustering of students into $K_2$ groups based on the similarity of their
transitioning behavior between the higher-level behaviors we can identify.

### 2017-06-18 sec-related-work.tex
Incorporate new citation from the editor.
Also shuffle around some figures to make the page layout prettier with
the now longer related work and model "derivation" sections.

A multi-layered approach to HMM modeling of sequence data has been
performed before in other domains. \citet{Zhang:2004:CVPR}, for example,
explored a two-layer HMM framework for modeling actions in meetings, but
their definition of ``two-layer'' differs from ours. In their formulation,
the ``lower-layer'' level is used to label audio-video action sequences
into basic events, and the ``upper-layer'' is used to label the output of
the lower-layer to discover higher-level office behavior abstractions. Our
formulation differs in that we do not feed the labeled sequence of the
lower level into the input of the higher level. Instead, our lower level is
actually treated using a non-hidden Markov model, and the higher level is
modeling transitions between the $K$ different non-hidden Markov models we
consider.

Our formulation is more closely related to the Hierarchical Hidden Markov
Model (HHMM) detailed in \citet{Fine:1998:ML}. Here, the ``layers'' are
modeled by having the hidden Markov model have two kinds of transitions.
Horizontal transitions move between states within a layer, where vertical
transitions move between different layers. At the bottom layer lie the
``production'' states, which output symbols according to some probability
distribution. Our specific model in this case can be modeled as a HHMM
where the horizontal transitions between nodes at the highest layer
(including self-loops) \emph{must} be immediately followed by a vertical
transition to the lower layer. The output probability distributions over
symbols in the lower layer ``production'' level are forced to emit only one
kind of symbol, and vertical transitions are only allowed into the original
higher-layer state that transitioned down into the lower-layer.

>>>>

A multi-layered approach to HMM modeling of sequence data has been
performed before in other domains. \citet{Zhang:2004:CVPR}, for example,
explored a two-layer HMM framework for modeling actions in meetings, but
their definition of ``two-layer'' differs from ours. In their formulation,
the ``lower-layer'' level is used to label audio-video action sequences
into basic events, and the ``upper-layer'' is used to label the output of
the lower-layer to discover higher-level office behavior abstractions.
\citet{Oliver:2004:CVIU} propose a similar layered HMM approach
for modeling office activity at multiple different levels of time
granularity. In their approach, each layer $L$ is represented as a ``bank''
of $K$ different HMMs that model sequences of some length $T_L$. At the
bottom layer ($L = 1$), the bank of $K$ HMMs corresponding to that layer is
run on some initial observation data, considering windows of observations
of length $T_1$. Then an output is generated by using the inferential
results of these $K$ HMMs to make a prediction: which of the $K$ HMMs was
most likely to produce that sequence of observations? This output is then
fed to the next layer of HMMs, which considers sequences of length $T_2$
and outputs prediction results as to which of the $K$ HMMs at layer two
were mostly likely to produce the sequence of outputs produced by the
previous layer, and so on.

Our formulation differs from both in that we do not feed the labeled
sequence of the lower level into the input of the higher level. Instead,
our lower level is actually treated using a non-hidden Markov model, and
the higher level is modeling transitions between the $K$ different
non-hidden Markov models we consider. The problem to be solved is similar
in that we wish to predict a ``label'' for a sequence of actions a student
takes as well as understand the transition behavior between those labels.
However, one of the consequences of modeling the lower layer using a
non-hidden Markov model instead of an HMM directly is that the meaning of
the $K$ different latent states can be more clearly captured by using our
proposed behavior representation. If we were to use an HMM as our first
layer, the behavior patterns (like from
Figure~\ref{fig:motivating-example}) would instead have nodes that
represent another set of latent states instead of being concrete actions
themselves. Each of these latent states would then be associated with some
other output distribution over the possible concrete actions to be
considered. In order to understand a single behavior pattern uncovered, one
would first have to understand the different output distributions for the
latent states in that pattern in order to understand the meaning of that
latent state. This further complicates the understanding of higher level
patterns because understanding the higher layer patterns requires
understanding the lower layer patterns. By taking a more restrictive view
on the first layer, we are able to produce a representation that can be
more readily interpreted due to the states in our first layer
representation having an immediately clear meaning (the concrete action
they represent).

The 2L-HMM model we propose is more closely related to the Hierarchical
Hidden Markov Model (HHMM) detailed in \citet{Fine:1998:ML}. Here, the
``layers'' are modeled by having the hidden Markov model have two kinds of
transitions.  Horizontal transitions move between states within a layer,
where vertical transitions move between different layers. At the bottom
layer lie the ``production'' states, which output symbols according to some
probability distribution. Our specific model in this case can be modeled as
a HHMM where the horizontal transitions between nodes at the highest layer
(including self-loops) \emph{must} be immediately followed by a vertical
transition to the lower layer. The output probability distributions over
symbols in the lower layer ``production'' level are forced to emit only one
kind of symbol, and vertical transitions are only allowed into the original
higher-layer state that transitioned down into the lower-layer.
